---
title: "Dang Gov 51 PSet 2"
author: "Tyler Dang"
format: pdf
editor: visual
---

# Part 1: The Original Study

## 1.1: Exploring the Experiment

### 1.1.1 - Load packages and data

```{r}
library(tidyverse)
library(dplyr)
library(broom)
library(knitr)
library(tinytex)

gay <- read.csv("data/raw/gay.csv")
```

### Q.1.1

```{r}
nrow(gay)
unique(gay$study)
unique(gay$treatment)
```

We can see, then, that there are 69,592 observations in the "gay" dataset. The "study" variable has two possible values: 1 and 2.

Meanwhile, the "treatment" variable has five possible values: No Contact, Same-Sex marriage Script by Gay Canvasser, Same-Sex Marriage Script by Straight Canvasser, Recycling Script by Gay Canvasser, or Recycling Script by Straight Canvasser.

### Q.1.2

```{r}
ssm_means <- gay %>%
  filter(study == 1 & wave == 1) %>%
  group_by(treatment) %>% 
  summarize(mean_ssm = mean(ssm))


```

We can see that the average support is very similar among the different treatment groups (for study one and wave one), but, Recycling Script by Gay Canvasser had the highest average Same Sex Marriage Rating attitude at 3.13.

### Q.1.3

```{r}
gay %>%
  filter(study == 1 & wave == 1) %>%
  group_by(treatment) %>% 
  summarize(sd_ssm = sd(ssm))

ssm_means
```

We can see that the standard deviations for the five treatment levels are very similar varying from 1.67 to 1.70. This is crucial for randomized testing as it supports the idea that if standard deviations and thus variances are similar among treatment groups, we can attribute the differences in treatment group means more to the treatment itself instead of confounding variables. In essence, the similarity of the standard deviations gives us more confidence that the differences in each treatment mean was because the treatment was imposed.

### Q.1.4

```{r}
ssm_means %>% 
  ggplot(mapping = aes(x = treatment, y = mean_ssm, fill = treatment)) +
  geom_col() +
  labs(title = "Mean SSM by Treatment Group (Study 1 and Wave 1)",
       x = "Treatment Group",
       y = "Mean SSM")+
  theme_minimal() +
  theme(axis.text.x = element_blank())
```

## 1.2 Checking Balance and Comparing Outcomes

### Q.1.5

```{r}
gay1.5 <- gay %>%
  filter(study == 1 & wave == 1) %>%
  filter(treatment %in% c("No Contact", "Same-Sex Marriage Script by Gay Canvasser")) %>% 
  group_by(treatment) %>% 
  summarise(mean_ssm = mean(ssm), 
            sd_ssm   = sd(ssm)) %>%
  mutate(std_diff_mean = 
           (mean_ssm[treatment == "Same-Sex Marriage Script by Gay Canvasser"] -
            mean_ssm[treatment == "No Contact"]) /
            sqrt((sd_ssm[treatment == "Same-Sex Marriage Script by Gay Canvasser"]^2 +
                  sd_ssm[treatment == "No Contact"]^2) / 2))

gay1.5
            
```

The standardized difference is −0.0105, which is extremely close to zero and far below the 0.25 threshold for imbalance. This indicates that the “Same-Sex Marriage Script by Gay Canvasser” and “No Contact” groups are essentially identical. The groups therefore appear very well balanced, consistent with successful randomization.

### Q.1.6

```{r}

gay1.6 <- gay %>%
  filter(study == 1 & wave == 2 & treatment %in% 
           c("Same-Sex Marriage Script by Gay Canvasser", "No Contact")) %>%
  group_by(treatment) %>%
  summarise( mean_ssm = mean(ssm))

gay1.6

gay1.6 %>%
  summarise(diff_means = 
            mean_ssm[treatment == "Same-Sex Marriage Script by Gay Canvasser"] -
            mean_ssm[treatment == "No Contact"])
```

The Wave 2 mean for No Contact is 3.0396, and for Same-Sex Marriage Script by Gay Canvasser it is 3.1395. The difference in means (gay canvasser − no contact) is 0.0999, or about 0.10 points.

Because the groups were essentially identical at baseline (standardized difference about −0.01), this Wave 2 difference suggests a causal effect of the treatment. Specifically, exposure to the same-sex marriage script delivered by a gay canvasser increased support for same-sex marriage by about 0.10 points relative to no contact.

# Part 2: The Forensic Discovery

## 2.1: The Baseline Distribution

### Q.2.1

```{r}
gay_r <- read_csv("data/raw/gayreshaped.csv")
ccap <- read_csv("data/raw/ccap2012.csv")
```

```{r}
nrow(gay_r)
nrow(ccap)

colnames(gay_r)
colnames(ccap)
```

We can see that there are 11948 observations in the "gay reshaped" dataset and 43998 in the ccap2012 dataset.

For gay reshaped, the column names are study, treatment, therm1, therm2, therm3, and therm4. meanwhile for the ccap2012 dataset, the column names are ..1, caseid, and gaytherm.

### Q.2.2

```{r}
gay_study1_stats <- gay_r %>%
  filter(study == 1) %>%
  summarise(mean = mean(therm1),
            median = median(therm1),
            sd = sd(therm1))

ccap_stats <- ccap %>%
  summarise(mean = mean(gaytherm, na.rm = TRUE),
            median = median(gaytherm, na.rm = TRUE),
            sd = sd(gaytherm, na.rm = TRUE))

bind_rows(GayR_Study1_Therm1 = gay_study1_stats,
          CCAP_GayTherm = ccap_stats, .id = "Dataset")
```

The two datasets look very similar. The means (58.38 vs. 58.71), medians (52 vs. 54), and standard deviations (28.50 vs. 29.37) are nearly identical, indicating very similar distributions of baseline feelings toward gay men and lesbians. This suggests that the Study 1 experimental sample closely resembles the broader CCAP 2012 sample in overall feeling toward gay people.

### Q.2.3

```{r}
gay_r2.3 <- gay_r %>%
  filter(study == 1) %>%
  select(therm1) %>%
  mutate(source = "LaCour Study 1", therm = therm1) %>%
  select(source, therm)

ccap_2.3 <- ccap %>%
  select(gaytherm) %>%
  filter(!is.na(gaytherm)) %>%
  mutate(source = "CCAP 2012", therm = gaytherm) %>%
  select(source, therm)

combined <- bind_rows(gay_r2.3, ccap_2.3)

ggplot(combined, aes(x = therm, fill = source)) +
  geom_histogram(breaks = seq(0, 100, by = 5)) +
  facet_wrap(~source) +
  labs(title = "Baseline Feeling Thermometer Comparison Between LaCour and CCAP",
       x = "Feeling Thermometer (0–100)",
       y = "Count") +
  theme(legend.position = "none")
```

Based on the two graphs, it is clear that the LaCour Study 1 and CCAP2012 baselines are from similar populations. If the two are truly randomly selected, the shape of the two distributions should show resemblance. As we can see, that is the case where there are three peaks: one near 0, one near 50 (which is the highest) and another near 100 (the second largest) on the feeling thermometer scale.

## 2.2: Re-test Reliability

### Q.2.4

```{r}
control <- gay_r %>%
  filter(treatment == "No Contact" & study %in% c(1, 2)) %>%
  group_by(study) %>%
  summarise(correlation = cor(therm1, therm2, use = "complete.obs"),
            sd_change = sd(therm2 - therm1, na.rm = TRUE))

control
```

Study 1 shows an extremely high re-test correlation (0.998), which is unrealisticly close to perfect stability. Its per-person change standard deviation (1.99) is also unusually small, indicating very little variation in responses between the waves. In contrast, Study 2’s correlation (0.973) is within the expected 0.95–0.97 range, and its larger change standard deviation (6.61) shows more realistic variation in responses. So, Study 2 has the more realistic pattern of test–retest reliability.

### Q.2.5

```{r}
control2.5 <- gay_r %>%
  filter(treatment == "No Contact", study %in% c(1, 2)) %>%
  select(study, therm1, therm2)

ggplot(control2.5, aes(x = therm1, y = therm2, color = study)) +
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0) +
  facet_wrap(~study, labeller = labeller(study = c(`1` = "Study 1",
                                                   `2` = "Study 2"))) +
  labs(title = "Test–Retest Feeling Thermometers (Control Group)",
       x = "Therm1", 
       y = "Therm2") +
  theme(legend.position = "none")
```

The Study 1 control group scatterplot shows points lying very closely to the 45-degree line, with many observations appearing identical across waves (especially at 0, 50, and 100). There is basically no variation around the line, which suggests nearly perfect test–retest connection. This is unrealistic for real longitudinal survey data where some measurement error and feeling fluctuation are expected.

Study 2 shows tight but clear scatter around the 45-degree line. Responses are strongly correlated but not identical, which reflects normal rounding, mood changes, and survey noise. Study 2 therefore looks much more like genuine longitudinal data.

## 2.3: Putting It All Together

### Q.2.6

From the testing that we have done, we have strong evidence that the data was fabricated instead of from real survey data. From Study 1, we found the baseline feeling thermometer distribution to be nearly identical to the CCAP2012 survey with means of 58.4 and 58.7 respectively and standard deviations of 28.5 and 29.4 respectively. This is unlikely if the samples were collected independently.

Moreover, we found the study 1 control re-test correlation to be .998 which is much higher than the typical .95-.97 range thus we further suspect fabrication as this suggests that respondents answered nearly identical from wave to wave.

Finally, the standard deviation of per-person change in study 1 was only 1.99 while in study 2 it was much larger at 6.61. We can argue for fabrication as the standard deviation in 1 suggest that there was very little individual fluctuation between waves.

One might produce data with these properties as it mimics an existing dataset thus ensuring realistic distributions, but there is little variation from the baseline producing patterns that seem consistent over time.

# Part 3: What Real Data Looks Like

## 3.1: Comparing Baselines

### Q.3.1

```{r}
gay_r_s2 <- gay_r %>%
  filter(study == 2)

nrow(gay_r_s2)

gay_r_s2 %>%
  group_by(treatment) %>%
  summarise(mean_therm1 = mean(therm1),
            sd_therm1 = sd(therm1),
            n = n())
```

There are 2441 observations in Study 2 of the gay reshaped data set. The mean therm1 response was 59.34 for the treatment compared to 56.89 for the control group, while the standard deviation for the treatment was 28.42 against 28.70 for the control.

### Q.3.2

```{r}
baseline3.2 <- gay_r %>%
  filter(study %in% c(1, 2)) %>%
  select(study, therm1) %>%
  mutate(source = ifelse(study == 1, "Study 1", "Study 2"))

ggplot(baseline3.2, aes(x = therm1, fill = source)) +
  geom_histogram(breaks = seq(0, 100, by = 5)) +
  facet_wrap(~source) +
  labs(title = "Baseline Feeling Thermometer Comparison",
       x = "Therm1 (0–100)",
       y = "Count") +
  theme(legend.position = "none")
```

From the graph we can see that the shape of the graph is very similar between study 1 and study 2. The biggest difference is the extremity of the graph (the count at each level is lower for study 2 than study 1). Also, study 1 has much more extreme spikes at its modes (0, 50, 100) while study 2 is much flatter. Going back to the idea of fabrication, the extreme spikes in study 1 look much more unrealistic.

## 3.2: Checking Balance and Comparing Outcomes in Study 2

### Q.3.3

```{r}
gay_r_S2_3.3 <- gay_r %>%
  filter(study == 2 & treatment %in% c("Same-Sex Marriage Script by Gay Canvasser",
                                       "No Contact")) %>% 
  group_by(treatment) %>%
  summarise(mean_therm1 = mean(therm1),
            sd_therm1 = sd(therm1)) %>% 
  mutate(std_diff_mean = 
           (mean_therm1[treatment == "Same-Sex Marriage Script by Gay Canvasser"] -
            mean_therm1[treatment == "No Contact"]) /
            sqrt((sd_therm1[treatment == "Same-Sex Marriage Script by Gay Canvasser"]^2 +
                  sd_therm1[treatment == "No Contact"]^2) / 2))

gay_r_S2_3.3
```

With a standardized difference of 0.051, which is very small and below the 0.25 threshold for imbalance, study 2 seems to be well balanced. We can also look at the baaseline means (57.89 vs. 59.35) and standard deviations (28.70 vs. 28.42) to see that these are also very similar across groups. So, the treatment and control groups in Study 2 appear well balanced at baseline, and thus probably conducted with true random assignment.

### Q.3.4

```{r}
gay_r_s2_3.4 <- gay_r %>%
  filter(study == 2 & treatment %in% c("Same-Sex Marriage Script by Gay Canvasser",
                                       "No Contact")) %>%
  group_by(treatment) %>%
  summarise(mean_therm2 = mean(therm2, na.rm = TRUE)) %>% 
  mutate(diff_means = mean_therm2[treatment == "Same-Sex Marriage Script by Gay Canvasser"] -
         mean_therm2[treatment == "No Contact"])

gay_r_s2_3.4
```

The post-period mean for the No Contact group is 57.64 while the mean for the Same-Sex Marriage Script by Gay Canvasser group is 61.20. The difference in post-period means (gay canvasser - no contact) is 3.57 points on the thermometer feeling scale.

## 3.3: Uncertainty Around the Difference

### Q.3.5

```{r}
gay_r_stas3.5 <- gay_r %>%
  filter(study == 2 & treatment %in% c("Same-Sex Marriage Script by Gay Canvasser",
                                       "No Contact")) %>%
  group_by(treatment) %>%
  summarise(mean_therm2 = mean(therm2, na.rm = TRUE),
            var_therm2  = var(therm2, na.rm = TRUE),
            n = n()) %>% 
  mutate(SE = sqrt(var_therm2[treatment == "Same-Sex Marriage Script by Gay Canvasser"] / 
                   n[treatment == "Same-Sex Marriage Script by Gay Canvasser"] + 
                   var_therm2[treatment == "No Contact"] /
                   n[treatment == "No Contact"]))

gay_r_stas3.5

diff_means <- gay_r_stas3.5$mean_therm2[gay_r_stas3.5$treatment == 
                                        "Same-Sex Marriage Script by Gay Canvasser"] - 
              gay_r_stas3.5$mean_therm2[gay_r_stas3.5$treatment == "No Contact"]

lower <- diff_means - 1.96 * gay_r_stas3.5$SE[1]
upper <- diff_means + 1.96 * gay_r_stas3.5$SE[1]

c(lower, upper)
```

Since our confidence interval from 1.29 to 5.84 does not include 0, had we conducted the sampling procedure many times, 95% of the resulting interval will capture the true difference in means. So, we can conclude that there is a statistically signnificant difference between the treatment and control groups. Canvassing on average positively affected the attitude towards gay populations by 1.29 to 5.84 points on the thermometer scale.

## 3.4: Tracking Attitudes Over Time

### Q.3.6

```{r}
s2_control <- gay_r %>%
  filter(study == 2 & treatment == "No Contact")

cor_therm2 <- cor(s2_control$therm1,
                  s2_control$therm2,
                  use = "complete.obs")

cor_therm3 <- cor(s2_control$therm1,
                  s2_control$therm3,
                  use = "complete.obs")

cor_therm4 <- cor(s2_control$therm1,
                  s2_control$therm4,
                  use = "complete.obs")

c(cor_therm2, cor_therm3, cor_therm4)
```

The correlations of the subsequent waves with the initial survey seem plausible. While they are high, each falls within the expected, typical .95-.97 range. While there is a dip in the third wave followed by an increase back in the fourth wave, these correlations are plausible, and the differences can be explained by sampling variability.

### Q.3.7

We can determine that Study 1 was fabricated due to the extreme differences from Study 2. In the first study, the re-test correlation between the first two waves was unrealistically high at .998 with a standard deviation of only 1.99. Study 2 had a much more realistic levels of these two with a correlation between waves falling within the .95-.97 range and a much larger standard deviation at 6.61.

The baseline distributions were also different. Study 1 was very similar to the CCAP national survey with means of 58.4 vs 58.7 and standard deviations of 28.5 vs 29.4. The nearly identical pattern of Survey 1 with the national survey is very unlikely given the smaller survey size. Survey 2 on the other hand displayed much more natural variation in these metrics. Thus, Study 1 seemed very unnatural and "clean" relative to the survey data.

Real treatment effects are smaller and messier than fabricated ones because real attitudes will fluctuate between rounds and people as respondents make rounding changes/errors, so the introduction of treatment will rarely provide such a large, consistent shift. Fabricated data may extremify the changes to more tidy effects.

Simple measurement tools aided in detecting the fraud by revealing the unlikely high stability with very small variation (sd) and very similar distributions to the national survey (means). Graphical analysis also shows that the "effects" of treatment were unnaturally extremified.

# Part 4: Formalizing the Evidence with Regression

## 4.1: Regression as a Fraud Detector

### Q.4.1

```{r}

fit_s1 <- lm(therm2 ~ therm1, data = gay_r,subset = study == 1)

tidy(fit_s1) %>%
  kable(digits = 3, caption = "Study 1: Regression Fit for Therm2 from Therm1")

summary(fit_s1)$r.squared
summary(fit_s1)$sigma
```

Our $\beta_1$ coefficient has a value of .991 which is the slope. It means that for a 1 point increase in the therm1 response, we would expect a corresponding .991 increase to the therm2 response.

The $R^2$ value is .98 allowing us to determine that 98% of the variance seen is explainable by therm1 for the regression model.

Our Residual Standard Error is 2.75 meaning that the average variance from the regression line is only 2.75 points on the thermometer scale.

### Q.4.2

```{r}
fit_s2 <- lm(therm2 ~ therm1, data = gay_r,subset = study == 2)

tidy(fit_s2) %>%
  kable(digits = 3, caption = "Study 2: Regression Fit for Therm2 from Therm1")

summary(fit_s2)$r.squared
summary(fit_s2)$sigma
```

Our $\beta_1$ coefficient has a value of .967 which is the slope. It means that for a 1 point increase in the therm1 response, we would expect a corresponding .967 increase to the therm2 response.

The $R^2$ value is .92 allowing us to determine that 98% of the variance seen is explainable by therm1 for the regression model.

Our Residual Standard Error is 7.96 meaning that the average variance from the regression line is only 2.75 points on the thermometer scale.

Compared to Study 1, we can see that Study 2 is much more realistic. In study 1, the $\beta_1$ coefficient is .991 which is very close to 1 indicating that the corresponding increase to therm2 from therm1 is nearly identical to that increase in therm1. Moreover, the $R^2$ value of .98 is also near 1 indicating that almost all of the variance in the regression model is due to therm1 (this eliminates an argument for large effects of confounding variables).

Meanwhile for Study 2, the $\beta_1$ coefficient, while large, is lower than in Study 1. The $R^2$ is also lower allowing for the possibility of confounding variables having effect in the model.

## 4.2: Comparing Groups with Regression

### Q.4.3

```{r}
gay_r_4.3 <- gay_r %>% 
  filter(study == 2 & treatment %in% c("Same-Sex Marriage Script by Gay Canvasser",
                                       "No Contact")) %>% 
  mutate(treated = case_when(treatment == 
                                     "Same-Sex Marriage Script by Gay Canvasser" ~ 1,
                                     treatment == "No Contact" ~ 0))

fit_4.2 <- lm(therm2 ~ treated, gay_r_4.3)

tidy(fit_4.2) %>%
  kable(digits = 3, caption = "Study 2: Regression Fit for Treatment on Control")

summary(fit_4.2)$r.squared
summary(fit_4.2)$sigma
```

Our $\beta_1$ coefficient has a value of 3.568 which is the slope. It means that for a 1 point increase in the therm1 response, we would expect a corresponding 3.568 increase to the therm2 response.

## 4.3: Controlling for Baseline Attitudes

### Q.4.4

### Q.4.5
